Dark Control:\\ A Unified Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-04-10 paper/paper.tex
style++

\begin{abstract}
The default mode network (DMN) is believed to subserve the human
baseline mental activity.
%
Many research streams converge to its likely role in evolutionarily adaptive
envisioning of mental scenarios to simulate the future.
Such functional roles become more likely due to
this brain network's highest energy consumption in the brain and
its intimate coupling with conscious awareness
\textbf{(XXX Guillaume: It escaladed quickly.)}.
%
The present concept paper proposes a \textit{process model}
in trying to explain \textit{how} the DMN may actually implement continuous
assessment and prediction of the environment to guide action choices.
DMN function is recast in the mathematical terms of
control theory and reinforcement learning based on Markov decision processes.
We argue that our formal account of DMN function
naturally accommodates
previously proposed cognitive accounts as special cases and
parsimoneously explains
experimental findings in animals and humans.
%
A process models for the neural implemention
of continuous future predictions probably realized in the DMN
could offer insight into the statistical mechanisms underlying human behavior.
\end{abstract}

@@

\section{Introduction}
%
When left unperturbed, the human brain is not at rest.
Rather, the brain continues to metabolize large quantities of
oxygen and glucose energy to maintain
potentially purposeful neuronal computation
in the absence of behavioral focus
\citep{kenet2003spontaneously, fiser2004small}.
The baseline energy demand is subject to surprisingly little modulations
when mentally processing environmental challenges
\citep{raichle2001pnas}. \textbf{(XXX Guillaume: What about the energy consumption compare to sleep? Danilo: I'd love to cite that
but I am not aware of such studies)}
This dark matter of brain physiology begs the question
of underlying biological purpose
\citep{raichle2006brain}.
What has early been described as the "stream of consciousness"
in psychology \citep{james1890principles}
found a potential neurobiological manifestation
in the so-called "default mode network" (DMN)
\citep{shul1997, raichle2001pnas}.
This set of some of the highest regions in
the human association cortex \citep{mesulam1998sensation, margulies2016situating}
consistently
increased in neural activity during
unfocused everyday mind wandering \citep{raichle2001pnas}.
The DMN may exert higher-order control on human behavior
following the notion of "a hierarchy of brain systems with
the DMN at the top and the salience and dorsal attention systems
at intermediate levels, above thalamic and unimodal sensory
cortex" \citep{carhart2010default}.
In the beginning of the 21st century,
brain imaging may have been first technique
to allow for the discovery of a unique brain network that subserves
baseline mental activities
\citep{raichle2001pnas, bzdok2015resting}.

@@

\subsection{Towards a computational account of DMN function}
We argue that the implication of the DMN in a
diversity of humans' most advanced cognitive processes
can be parsimoniously recast as prediction error minimization
based on pervasive probabilistic simulations,
thus maximizing reward outcome across multiple time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo sampling
\citep{tenenbaum2011grow}.
In this context,
\textit{control} refers to the influences that an agent exerts when interacting
with the environment to encourage preferred states.
Even (necessarily imperfect) memories, random day-time mind-wandering, and dreams during sleep
may provide meaningful building blocks to iteratively improve
the predictive DMN machinery to optimize the behavior of the organism.
%
In short, it has been proposed earlier that
the human brain's energy budget is largely dedicated to
``the development and maintenance of [a]
probabilistic model of anticipated events''
\citep{raichle2005intrinsic}.
Numerous empirical evidence from
neuroscientific experiments have invigorated this idea \citep{kording2004bayesian, fiser2004small}.
This paper proposes a process model that satisfies this contention.

\section{Known neurobiological properties of the default mode network}
We begin by a neurobiological deconstruction of the DMN
based on experimental findings in neuroscience. This walkthrough across individual network nodes will describe their mechanistic relevance, paving the way for their functional interpretation within our computational account in the next section.

\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_overview_DMN.pdf}
  \caption{Neurobiological overview of the DMN ... \textbf{Please add a figure caption}}
  \label{fig:neurobiological_overview_DMN}
\end{figure}
%
\subsection{The posteromedial cortex: global monitoring and information integration}
The midline structures of the DMN,
including the posteromedial cortex (PMC) and
the medial prefrontal cortex (mPFC),
are responsible for the highest metabolic turn-over
of glucose energy consumption \citep{raichle2001pnas}.
These metabolic characteristics go hand-in-hand with
brain-imaging analyses \citep{andrews2010} in
suggesting that the PMC and mPFC
potentially represent the functional backbone of the DMN.

The PMC myelinates relatively late during postnatal development in monkeys
\citep{goldman1987development}, which is generally considered to
be a sign of evolutionary sophistication \citep{flechsig1920}.
Physiological and disturbed metabolic fluctuations in the
human PMC have been repeatedly related to
phenomena of changed conscious awareness \citep{cavanna2006precuneus},
including anesthesia, sleep, and coma.
%
Indeed,
the PMC has long been speculated to reflect constant computation of
environmental statistics and its internal representation
as an "inner minds eye" \citep{cavanna2006precuneus, leech_pcc2014}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that results from damage in the parietal cortex
\citep{balint1909seelenlahmung}.
Patients are plagued by an
inability to bind various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
This can be viewed as a high-level impairment in the gathering
of information about alternative objects (i.e., exploration) as well as
leveraging these environmental opportunities (i.e., exploitation).
Congruently,
the human PMC was coupled in two functional connectivity modalities
with the amygdala, involved in significance evaluation, and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological significance assessment.

The putative role of the PMC in continuous integration of environmental relevance
and ensuing top-level control of action on the environment is supported
by a number of neuroscience experiments.
Electrophysiological recordings in animals implicated the PMC in
strategic decision making \citep{pearson2009neurons},
risk assessment \citep{mccoy2005risk},
and outcome-contingent behavioral modulation \citep{hayden2009electrophysiological},
while its retrosplenial portion was
more specifically implicated in approach-avoidance behavior
\citep{vann2009does}.
Neuron spiking activity in the PMC allowed distinguishing
whether a monkey would persue an exploratory or exploitative
behavioral strategy in a complex food foraging task \citep{pearson2009neurons}.
Further, single-cell recordings in the monkey PMC
demonstrated this brain region's sensitivity to
subjective target utility \citep{mccoy2005risk} and integration
across individual decision-making instances \citep{pearson2009neurons}.
This DMN node encoded the
preference of or aversion to options with uncertain reward outcomes
and its spiking activity was more associated with
subjectively perceived relevance of a chosen object
than by its factual value, based on an internal currency of value.
In fact, direct stimulation of PMC neurons
promoted exploratory action towards options
with unsafe reward outcomes that were previously shunned.
Graded changes in firing rates of PMC neurons
indicated changing choices in upcoming trials and their neural patterns were
distinct from spike firings that indicated choosing either option.
Similarly in humans,
the DMN has been shown to gather and integrate information
over different paragraphs in an fMRI study on auditory narratives
\citep{simony2016dynamic}.

@@

Consequently, the PMC may contribute to overall DMN function
by monitoring the subjective outcomes
of possible decisions and integrating that information
with memory, perspective frames, and
reward schedules into higher-level strategies.
Perceived value, that differs across individuals, may update
statistical assessment of the environment
to predict delayed reward opportunities in the future.
In doing so, the PMC continuously adapts to changes
in both the external environment and its internal representation
to enable strategic behavior in volatile environments.

@@

Generally,
patients with neurological lesions in the prefrontal cortex
are known to struggle with
adaption to novel situations and stimuli
\citep{stuss1986frontal}.
The dmPFC may subserve inference, representation, and assessment
of one's own and other individuals' action considerations.
Specifically, neural activity in the human dmPFC
reflected expectations about other peoples' actions and
errors in these predictions.
Neural activity in the dmPFC indeed explained the proficiency decline
of inferring other peoples' thoughts in the elderly \citep{moran2012social}.
Some dmPFC neurons in macaque monkeys exhibited a preference
for processing others', rather than own, behavior
with fine-grained adjustment of contextual circumstances \citep{yoshida2010neural}.
In fact, the topographically neighboring dorsal anterior cingulate cortex
has also been linked to computing values and efforts of
persisting a behavioral plan versus switching the
environmental context in several lesion studies \citep{kolling2016value}.
%
Such highly abstract neural computations necessarily rely on the
generation of probabilistic internal information drawing from
episodic memory retrieval, generative construction processes,
and explicit knowledge of the external world.
%
According to computational neuroimaging experiments,
dmPFC activity preferentially models stimulus-value associations of
possible actions that are not actually executed,
whereas the vmPFC activity models value in the environment that leads performed actions
\citep{nicolle2012agent}.

Comparing to the dorsomedial prefrontal DMN,
the ventromedial portion of the prefrontal cortex (vmPFC) probably subserves
less ambiguous subjective-value-related evaluative processes and
reward-informed risk estimation of self-relevant environmental stimuli.
This DMN node is more closely associated with
orchestrating adaptive behavior by bottom-up-driven
processing of “what matters now”,
probably drawing on sophisticated value representations
\citep{doherty2015structure}.
Quantitative lesion findings across 344 human individuals confirmed
a substantial impairment in value-based decision making
\citep{glascher2012lesion}.
Indeed,
the vmPFC is preferentially connected with reward-related and limbic areas.
The vmPFC is well known to have monosynaptical connections
with the NAc
in axonal tracing studies in monkeys \citep{haber1995orbital}.
Congruently, the gray-matter volume of the vmPFC and NAc
correlated with indices of value-guided behavior and reward attitudes
in humans
\citep{lebreton2009automatic}.
NAc activity is thought to reflect reward prediction signals from dopaminergic pathways
\citep{schultz1998predictive}
that not only modulate motivated behavior towards basic survival needs but also
subserve RL in humans more broadly \citep{doherty2015structure}.
This is consistent with diffusion MRI tractography in humans and monkeys
\citep{croxson2005quantitative} that
quantified the NAc to
be substantially more likely connected to the vmPFC than dmPFC in both species.
Two functional connectivity modalities in humans strongly connected
the vmPFC with the NAc, hippocampus (HC),
and PMC.
%
The vmPFC is often proposed to be involved in (external) emotional
reactions and own (visceral) arousal.
Real or imaginged bodily states could be mapped in the vmPFC
as a bioregulatory disposition governing cognition
and decision making \citep{damasio1996somatic}
In neuroeconomic studies of human decision making,
the vmPFC consistently reflects an individual’s subjective
valuation
\citep{behrens2008associative}.
This may be why performance within and across participants
was related to state encoding in vmPFC.
Such a ``cognitive map'' of the action space was argued to encode
the current task state even when states are unobservable from sensory input,
which was shown to be critical for behavior \citep{Schuck2016}.
Additionally,
independent whole-brain analyses from structural
neuroimaging studies related the gray-matter volume of the vmPFC,
more consistently than any other
brain region, to indices of
social competence and social network complexity, which are probably
among the most complicated decision that humans and monkeys take
\citep{behrens2009computation}.

@@

RL is a problem-solving technique in which, through interactions with an external environment,
an agent learns to reach goals, optimize reward signals, or minimize costs
in an iterative trial-by-error fashion.
At a given moment, each taken \textit{action} $a$ triggers a change in the
\textit{state} of the environment
$s \rightarrow s'$, accompanied by environmental feed signals as \textit{reward}
$r = r(s, a)$ (or \textit{punishment},
since it can be negative) collected by the
agent.
In this view, the environment is partially controlled by
the action of the agent and the reward can be thought
of as satisfaction -- or punishment -- accompanying the execution of
an action.
Action may be delayed to achieve substantial increases in expected reward
that can grow with time.
The environment is generally taken as stochastic,
that is, changing in random ways. In addition,
it is only \textit{partially observed},
in the sense that only part of the current state is observed by
the agent \citep{starkweather2017dopamine}.
The unpredictability of the environment
is realistic in a computational model which sets out
to explain DMN function.
% XXX: the following passage is already in the introduction
% Indeed, it is known that the more the external world is predictable,
% the more mental activity becomes detached from the actual sensory inputs
% \citep{antrobus1966studies, pope1978regulation},
% the more stimulus-independent thoughts occur, and
% the higher the neural activity in the DMN.
% Conversely, the more the currently executed task is unknown and unpracticed,
% the lower the DMN activity
% \citep{filler1973daydreaming, teasdale1995stimulus}.
% %
% These ``offline'' processes may however contribute to
% optimizing control of the organism.
% Such a framework of DMN funtion based on reinforcement learning
% can naturally embed human behavior
% in the tension between exploitative action with immediate gains and
% explorative action with longer-term reward schedules
% \citep{dayan2008decision}.
A framework of DMN funtion based on RL
can naturally embed human behavior
in the tension between exploitative action with immediate gains and
explorative action with longer-term reward schedules
\citep{dayan2008decision}.
DMN implication in a diversity of
particularly advanced cognitive processes
can be parsimoneously explained as probabilistic simulations in form of mental scenes coupled with prediction error minimization
to calibrate action portfolios and
maximize reward outcome at different time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling
\citep{tenenbaum2011grow}.

@@

More specifically,
environmental stimuli especially important for humans are frequently of
social nature. This is probably unsurprising
given that 
the complexity of the social systems
is a likely human-defining property
\citep{tomasello2009cultural}.
According to the ``social brain hypothesis'',
the human brain has especially been shaped for
forming and maintaining increasingly complex
social systems,
which allows solving ecological problems by social relationships
\citep{whiten1988machiavellian}.
Indeed, social topics amount to roughly
two thirds of human everyday communication \citep{dunbar1997human},
while
mind-wandering at daytime and dreams during sleep
are rich in stories about people and
the complex relationships between them.
%
In line with this, the DMN has been proposed to reflect
continuous processing of social information as a
physiological baseline of human brain function
\citep{schilbach2008minds}. This view was later challenged by observing
analogues of the DMN in monkeys \citep{mantini2011default}
and rats \citep{lu2012rat}, both species with
supposedly less advanced social-cognitive capacities.
Rather than functional specificity for processing social information,
the present formal account can parsimoneously incorporate
the dominance of social content in
human mental activity 
as high value function estimates for information about humans
\citep{baker2009action, kampe2001psychology, krienen2010clan}.
The DMN may thus exert control on the human agent in a way that prioritizes
appraisal of and action towards social contexts,
without entirely neglecting environmental cues of the physical world.
In sum,
our account on the DMN directly implies
its previously proposed ``sentinel'' function
of monitoring the environment for salient information
in general and
inherently incorporates the importance of social environmental cues
as an important special case.

>>>>

\begin{abstract}
The default mode network (DMN) is believed to subserve the human
baseline mental activity.
%
Many research streams converge to its likely role in evolutionarily adaptive
envisioning of mental scenarios to simulate the future.
Such functional roles become more likely due to
this brain network's highest energy consumption in the brain and
its intimate coupling with conscious awareness
\textbf{(XXX Guillaume: It escaladed quickly.)}.
%
The present concept paper proposes a \textit{process model}
in trying to explain \textit{how} the DMN may actually implement continuous
assessment and prediction of the environment to guide action choices.
DMN function is recast in the mathematical terms of
control theory and reinforcement learning based on Markov decision processes.
We argue that our formal account of DMN function
naturally accommodates
previously proposed cognitive accounts as special cases and
parsimoniously explains
experimental findings in animals and humans.
%
A process models for the neural implemention
of continuous future predictions probably realized in the DMN
could offer insight into the statistical mechanisms underlying human behavior.
\end{abstract}

@@

\section{Introduction}
%
When left unperturbed, the human brain is not at rest.
Rather, the brain continues to metabolize large quantities of
oxygen and glucose energy to maintain
neuronal computation
in the absence of behavioral focus
\citep{kenet2003spontaneously, fiser2004small}.
The baseline energy demand is subject to surprisingly little modulations
when mentally processing environmental challenges
\citep{raichle2001pnas}. \textbf{(XXX Guillaume: What about the energy consumption compare to sleep? Danilo: I'd love to cite that
but I am not aware of such studies)}
This dark matter of brain physiology begs the question
of underlying biological purpose
\citep{raichle2006brain}.
What has early been described as the "stream of consciousness"
in psychology \citep{james1890principles}
found a potential neurobiological manifestation
in the so-called "default mode network" (DMN)
\citep{shul1997, raichle2001pnas}.
This set of some of the highest regions in
the human association cortex \citep{mesulam1998sensation, margulies2016situating}
consistently
increased in neural activity during
unfocused everyday mind wandering \citep{raichle2001pnas}.
The DMN may exert higher-order control on human behavior
following the notion of "a hierarchy of brain systems with
the DMN at the top and the salience and dorsal attention systems
at intermediate levels, above thalamic and unimodal sensory
cortex" \citep{carhart2010default}.
In the beginning of the 21st century,
brain imaging may have been the first technique
to allow for the discovery of a unique brain network that subserves
baseline mental activities
\citep{raichle2001pnas, bzdok2015resting}.

@@

\subsection{Towards a formal account of DMN function}
We argue that the implication of the DMN in a
diversity of humans' most advanced cognitive processes
can be parsimoniously recast as prediction error minimization
based on pervasive probabilistic simulations,
thus maximizing reward outcome across multiple time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo sampling
\citep{tenenbaum2011grow}.
In this context,
\textit{control} refers to the influences that an agent exerts when interacting
with the environment to encourage preferred states.
Even (necessarily imperfect) memories, random day-time mind-wandering, and dreams during sleep
may provide meaningful building blocks to iteratively improve
the predictive DMN machinery to optimize the behavior of the organism.
%
In short, it has been proposed earlier that
the human brain's energy budget is largely dedicated to
``the development and maintenance of [a]
probabilistic model of anticipated events''
\citep{raichle2005intrinsic}.
This idea has been invigorated by
empirical evidence from
neuroscientific experiments \citep{kording2004bayesian, fiser2004small}.
The present paper proposes a
process model that satisfies this contention.

\section{Known neurobiological properties of the default mode network}
We begin by a neurobiological deconstruction of the DMN
based on experimental findings in the neuroscience literature.
This walkthrough across network nodes will
outline their individual functional profiles,
paving the way for their functional interpretation within our computational account
described in the next section.

\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_overview_DMN.pdf}
  \caption{Neurobiological overview of the DMN ... \textbf{Please add a figure caption}}
  \label{fig:neurobiological_overview_DMN}
\end{figure}
%
\subsection{The posteromedial cortex: global monitoring and information integration}
The midline structures of the DMN,
including the posteromedial cortex (PMC) and
the medial prefrontal cortex (mPFC),
are responsible for the highest turn-over
of glucose energy consumption \citep{raichle2001pnas}.
These metabolic characteristics go hand-in-hand with
brain-imaging analyses \citep{andrews2010} in
suggesting that the PMC and mPFC
potentially represent the functional backbone of the DMN.

The PMC myelinates relatively late during postnatal development in monkeys
\citep{goldman1987development}, which is generally considered to
be a sign of evolutionary sophistication \citep{flechsig1920}.
Normal and disturbed metabolic fluctuations in the
human PMC have been repeatedly related to
phenomena of changed conscious awareness \citep{cavanna2006precuneus},
including anesthesia, sleep, and coma.
%
Indeed,
the PMC has long been speculated to reflect constant computation of
environmental statistics and its internal representation
as an inner "mind's eye" \citep{cavanna2006precuneus, leech_pcc2014}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that results from damage in the parietal cortex
\citep{balint1909seelenlahmung}.
Patients are plagued by an
the inability to bind various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
This can be viewed as a high-level impairment in gathering
information about alternative objects (i.e., exploration) as well as
leveraging these environmental opportunities (i.e., exploitation).
Congruently,
the human PMC was coupled in two functional connectivity analyses
with the amygdala, involved in significance evaluation, and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological significance assessment.

The putative role of the PMC in continuous integration of
relevant environmental information
and ensuing top-level control of action on the environment is supported
by a number of neuroscience experiments.
Electrophysiological recordings in animals implicated the PMC in
strategic decision making \citep{pearson2009neurons},
risk assessment \citep{mccoy2005risk},
and outcome-contingent behavioral modulation \citep{hayden2009electrophysiological},
while its retrosplenial portion was
repeatedly implicated in approach-avoidance behavior
\citep{vann2009does}.
Neuron spiking activity in the PMC allowed distinguishing
whether a monkey would persue an exploratory or exploitative
behavioral strategy during food foraging \citep{pearson2009neurons}.
Further, single-cell recordings in the monkey PMC
demonstrated this brain region's sensitivity to
subjective target utility \citep{mccoy2005risk} and integration
across individual decision-making instances \citep{pearson2009neurons}.
This DMN node encoded the
preference of or aversion to options with uncertain reward outcomes
and its spiking activity was more associated with
subjectively perceived relevance of a chosen object
than by its factual value,
based on an internal currency of value \citep{mccoy2005risk}.
In fact, direct stimulation of PMC neurons
promoted exploratory action towards actions
that were previously shunned \citep{hayden2008stim}.
Graded changes in firing rates of PMC neurons
indicated changing choices in upcoming trials and their neural patterns were
distinct from spike firings that indicated choosing either option.
Similarly in humans,
the DMN has been shown to gather and integrate information
over different paragraphs in an fMRI study on auditory narratives
\citep{simony2016dynamic}.

@@

Consequently, the PMC may contribute to overall DMN function
by monitoring the subjective outcomes
of possible decisions and integrating that information
with memory, perspective frames, and
reward schedules into higher-level strategies.
Perceived value, that differs across individuals, may update
statistical assessment of the environment
to predict delayed reward opportunities in the future.
In doing so, the PMC may continuously adapt to changes
in both the external environment and its internal representation
to enable strategic behavior in volatile environments.

@@

Generally,
patients with neurological lesions in the prefrontal cortex
are known to struggle with
adaption to novel situations and stimuli
\citep{stuss1986frontal}.
The dmPFC may subserve inference, representation, and assessment
of one's own and other individuals' action considerations.
Specifically, neural activity in the human dmPFC
reflected expectations about other peoples' actions and
errors in these predictions.
Neural activity in the dmPFC indeed explained the proficiency decline
of inferring other peoples' thoughts in the elderly \citep{moran2012social}.
Some dmPFC neurons in macaque monkeys exhibited a preference
for processing others', rather than own, behavior
with fine-grained adjustment of contextual circumstances \citep{yoshida2010neural}.
In fact, the topographically neighboring dorsal anterior cingulate cortex
has also been linked to computing values and efforts of
persisting a behavioral plan versus switching the
environmental context in several lesion studies \citep{kolling2016value}.
%
Such highly abstract neural computations necessarily rely on the
generation of probabilistic internal information drawing from
episodic memory retrieval, generative construction processes,
and explicit knowledge of the external world.
%
According to computational neuroimaging experiments,
dmPFC activity preferentially models stimulus-value associations of
possible actions that are not actually executed,
whereas vmPFC activity models value in the environment that leads performed actions
\citep{nicolle2012agent}.

Comparing to the dorsomedial prefrontal DMN,
the ventromedial portion of the prefrontal cortex (vmPFC) probably subserves
subjective-value-related evaluative processes and
risk estimation of self-relevant environmental stimuli.
This DMN node is more closely associated with
orchestrating adaptive behavior by bottom-up-driven
processing of “what matters now”,
probably drawing on sophisticated value representations
\citep{doherty2015structure}.
Quantitative lesion findings across 344 human individuals confirmed
a substantial impairment in value-based decision making
\citep{glascher2012lesion}.
Indeed,
the vmPFC is preferentially connected with reward-related and limbic areas.
The vmPFC is well known to have monosynaptical connections
with the NAc
in axonal tracing studies in monkeys \citep{haber1995orbital}.
Congruently, the gray-matter volume of the vmPFC and NAc
correlated with indices of value-guided behavior and reward attitudes
in humans
\citep{lebreton2009automatic}.
NAc activity is thought to reflect reward prediction signals from dopaminergic pathways
\citep{schultz1998predictive}
that not only modulate motivated behavior towards basic survival needs but also
subserve RL in humans more broadly \citep{doherty2015structure}.
This is consistent with diffusion MRI tractography in humans and monkeys
\citep{croxson2005quantitative} that
quantified the NAc to
be substantially more likely connected to the vmPFC than dmPFC in both species.
Two functional connectivity analyses in humans strongly connected
the vmPFC with the NAc, hippocampus (HC),
and PMC.
%
The vmPFC is often proposed to be involved in (external) emotional
reactions and own (visceral) arousal.
Real or imagined bodily states could be mapped in the vmPFC
as a bioregulatory disposition governing cognition
and decision making \citep{damasio1996somatic}
In neuroeconomic studies of human decision making,
the vmPFC consistently reflects an individual’s subjective
valuation
\citep{behrens2008associative}.
This may be why performance within and across participants
was related to state encoding in the vmPFC.
Such a ``cognitive map'' of the action space was argued to encode
the current task state even when states are unobservable from sensory input,
which was shown to be critical for behavior \citep{Schuck2016}.
% Additionally,
% independent whole-brain analyses from structural
% neuroimaging studies related the gray-matter volume of the vmPFC,
% more consistently than any other
% brain region, to indices of
% social competence and social network complexity, which are probably
% among the most complicated decision that humans and monkeys take
% \citep{behrens2009computation}.

@@

RL is a problem-solving technique in which, through interactions with an external environment,
an agent learns to reach goals, optimize reward signals, or minimize costs
in an iterative trial-by-error fashion.
At a given moment, each taken \textit{action} $a$ triggers a change in the
\textit{state} of the environment
$s \rightarrow s'$, accompanied by environmental feed signals as \textit{reward}
$r = r(s, a)$ (or \textit{punishment},
since it can be negative) collected by the
agent.
In this view, the environment is partially controlled by
the action of the agent and the reward can be thought
of as satisfaction -- or punishment -- accompanying the execution of
an action.
Action may be delayed to achieve substantial increases in expected reward
that can grow with time.
The environment is generally taken as stochastic,
that is, changing in random ways. In addition,
it is only \textit{partially observed},
in the sense that only part of the current state is observed by
the agent \citep{starkweather2017dopamine}.
The unpredictability of the environment
is realistic in a computational model which sets out
to explain DMN function.
% XXX: the following passage is already in the introduction
% Indeed, it is known that the more the external world is predictable,
% the more mental activity becomes detached from the actual sensory inputs
% \citep{antrobus1966studies, pope1978regulation},
% the more stimulus-independent thoughts occur, and
% the higher the neural activity in the DMN.
% Conversely, the more the currently executed task is unknown and unpracticed,
% the lower the DMN activity
% \citep{filler1973daydreaming, teasdale1995stimulus}.
% %
% These ``offline'' processes may however contribute to
% optimizing control of the organism.
% Such a framework of DMN funtion based on reinforcement learning
% can naturally embed human behavior
% in the tension between exploitative action with immediate gains and
% explorative action with longer-term reward schedules
% \citep{dayan2008decision}.
A framework of DMN funtion based on RL
can naturally embed human behavior
in the tension between exploitative action with immediate gains and
explorative action with longer-term reward schedules
\citep{dayan2008decision}.
DMN implication in a diversity of
particularly advanced cognitive processes
can be parsimoniously explained as probabilistic simulations in form of mental scenes coupled with prediction error minimization
to calibrate action portfolios and
maximize reward outcome at different time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling
\citep{tenenbaum2011grow}.

@@

More specifically,
environmental stimuli especially important for humans are frequently of
social nature. This is probably unsurprising
given that 
the complexity of the social systems
is a likely human-defining property
\citep{tomasello2009cultural}.
According to the ``social brain hypothesis'',
the human brain has especially been shaped for
forming and maintaining increasingly complex
social systems,
which allows solving ecological problems by social relationships
\citep{whiten1988machiavellian}.
Indeed, social topics amount to roughly
two thirds of human everyday communication \citep{dunbar1997human},
while
mind-wandering at daytime and dreams during sleep
are rich in stories about people and
the complex relationships between them.
%
In line with this, the DMN has been proposed to reflect
continuous processing of social information as a
physiological baseline of human brain function
\citep{schilbach2008minds}. This view was later challenged by observing
analogues of the DMN in monkeys \citep{mantini2011default}
and rats \citep{lu2012rat}, both species with
supposedly less advanced social-cognitive capacities.
Rather than functional specificity for processing social information,
the present formal account can parsimoniously incorporate
the dominance of social content in
human mental activity 
as high value function estimates for information about humans
\citep{baker2009action, kampe2001psychology, krienen2010clan}.
The DMN may thus exert control on the human agent in a way that prioritizes
appraisal of and action towards social contexts,
without entirely neglecting environmental cues of the physical world.
In sum,
our account on the DMN directly implies
its previously proposed ``sentinel'' function
of monitoring the environment for salient information
in general and
inherently incorporates the importance of social environmental cues
as an important special case.
